Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST

### 2018-12-01 paper/tex/methods.tex
fix huge spacing for two equations

We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes. Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space. Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space. We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:

\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}

\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.

@@

\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise). The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$. The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:

\begin{align}
f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
\end{align}
where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively. Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy. The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.

>>>>

We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes. Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space. Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space. We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}
\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.

@@

\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise). The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$. The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
\begin{equation}
f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
\end{equation}
where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively. Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy. The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
